Why there are more species in several small patches vs few large patches:

### 2024-08-02 ms_SLOSS.Rmd
New simulations with deforestation-at-time=600

1. As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it. The policies of many conservation agencies are still prioritizing the preservation of a single large habitat area (SL) vs. several small areas of the same size (SS), despite empirical support favouring SS. However, to date, most theories regarding this issue were based on single-species models.

@@

The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021]. Despite the long-standing discussion, there remains a notable scarcity of multi-species models explicitly designed to validate fragmentation theories. However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].

@@

<!-- In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).-->

@@

For all simulations, we used a metacommunity of 100 species. The simulations start with the colonization of an empty habitat area. After 600 time steps, a fixed amount of 60% of habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss. <!-- why we don't use completely random degradation patterns -->

>>>>

1. As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it. The policies of many conservation agencies are still prioritizing the preservation of a single large habitat area (SL) vs. several small areas of the same size (SS), despite empirical support favouring SS. However, to date, while many studies have explored this issue using model communities, the results are mixed. In competitive community models, fragmentation has a negative effect at low habitat amounts, whereas neutral community models suggest positive effects.

@@

The SLOSS debate has been approached from several theoretical frameworks, but many cases have been based based on single-species models [@Fahrig2021]. Despite the long-standing discussion, there remains a notable scarcity of multi-species models explicitly designed to validate fragmentation theories. However, recent studies using spatially explicit simulations of competitive communities have revealed that while high fragmentation leads to more local extinctions due to reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020]. Conversely, studies based on neutral models have shown that fragmentation can have a positive effect on species richness, particularly when the habitat amount is low [@Smith2024]. Additionally, spatially implicit models with competition-colonization trade-offs result in nonlinear oscillatory responses to both habitat loss and fragmentation [@Zhang2023].

@@

@@

For all simulations, we used a metacommunity of 100 species. The simulations begin with the colonization of an empty habitat area. After 600 time steps, a fixed amount of habitat (40%, 60%, or 80%) is subsequently subjected to destruction, employing either regular or random block degradation to mimic habitat fragmentation due to habitat loss. The results were collected after 1500 time steps. <!-- why we don't use completely random degradation patterns -->
